\chapter*{Acknowledgements}

A lot of people contributed to this thesis and I will explicitly thank everyone, in no particular order.

I thank my academic supervisors, Professors André Restivo and Pedro Ribeiro, for their valuable insights on how to conduct a research project, namely a master thesis, and how to structure this final document. I thank them for the multiple times where they advised me on which direction to take and the multiple great ideas they contributed with. Above all, I thank them for always putting my interests first and ensuring I had the necessary conditions to research, develop and write my thesis.

I thank my company supervisors, Marco Sampaio and Pedro Silva, for the advice they gave me throughout this thesis which made it possible for me to develop this project, step by step. I thank them for the hours they spent discussing with me the next steps to take, experiments to conduct, new methods to try and for the remaining hours they spent analyzing the results with me. Finally, I thank them for the many hours spent reviewing this document.

I thank the entire Feedzai Systems Research team, namely, João Oliveirinha, Pedro Silva and Sofia Gomes for their daily support and suggestions. I thank them for the technical tips they gave me to surpass certain tasks that at first seemed impossible. 

Last but certainly not least, I thank my family, friends and girlfriend for keeping me sane while I worked on this thesis during the COVID-19 lockdown.

\vspace{10mm}
\flushleft{João Dias Conde Azevedo}